---
title: "DESeq2/edgeR Analysis"
output: html_notebook
---
This protocol is for executing DESeq2/edgeR after properly preparing the Seurat 
object file (SOF). As a prerequisite, make sure that the cell types (e.g., 
Cluster ID) and experimental groups are as a factor in the SOF. Example code is 
listed below.
```{r}
aggr_seurat@meta.data$Cluster_ID <- as.factor(aggr_seurat@meta.data$Cluster_ID)
aggr_seurat@meta.data$Groups <- as.factor(aggr_seurat@meta.data$Groups)
```
Make sure that the active identity is set to the experimental group column in 
the metadata. Additionally, ensure that the replicate (i.e., animal ID)
information is included in the metadata. 
```{r}
aggr_seurat <- SetIdent(aggr_seurat, value = aggr_seurat@meta.data$Groups)

metadata <- read.csv("seurat_aggr_metadata.csv")
Animal_ID <- metadata[1:165363,6]
aggr_seurat$Replicates <- Animal_ID
```
If needed, you can extract the metadata file as a .csv to edit or input 
additional information in column. This .csv file can then be imported into R to
adjust the metadata from the SOF.
```{r}
write.csv(aggr_seurat@meta.data, "./seurat_metadata.csv")
```
Once the SOF is finalized, you can save it into an RDS or Rdata file.
```{r}
save(aggr_seurat, file = "Aging_Seurat.Rdata")
saveRDS(aggr_seurat, file = "Aging_Seurat.RDS")
```
Next, install the Libra package, and subset the SOF to include the two groups
you will use in your comparisons. Make sure that the control group is listed
SECOND. Otherwise, the data will be "backwards." Alternatively, you can relevel
the data using relevel() and ensuring that ref = "treatment group".
``` {r}
library(Libra)
library(DESeq2)

female_aging <- subset(aggr_seurat, idents = c("Female-AV", "Female-YV"))
male_aging <- subset(aggr_seurat, idents = c("Male-AV", "Male-YV"))
male_treatment <- subset(aggr_seurat, idents = c("Male-AM", "Male-AV"))
 
DE_female_aging <- run_de(female_aging, cell_type = "Cluster_ID", label = 
                            "Groups", replicate = "Replicates", de_family = 
                            'pseudobulk', de_method = 'edgeR', de_type = 'LRT')

DE_male_aging <- run_de(male_aging, cell_type = "Cluster_ID", label = "Groups",
                          replicate = "Replicates", de_family = 'pseudobulk',
                          de_method = 'edgeR', de_type = 'LRT')

DE_male_treatment <- run_de(male_treatment, cell_type = "Cluster_ID", label = 
                              "Groups", replicate = "Replicates", de_family = 
                              'pseudobulk', de_method = 'edgeR', de_type = 
                              'LRT')

write.csv(DE_female_aging, "DE Female Aging edgeR.csv", row.names = FALSE, 
          quote = FALSE)
write.csv(DE_male_aging, "DE Male Aging edgeR.csv", row.names = FALSE, 
          quote = FALSE)
write.csv(DE_male_treatment, "DE Male Treatment edgeR.csv", row.names = FALSE, 
          quote = FALSE)
```
Below is some additional information regarding SOFs and miscellaneous analyses
that can be performed after retrieving the differential expression data. 
```{r}
Idents(aggr_seurat) # This tells you the levels of the active identity.

table(aggr_seurat@meta.data$Groups,aggr_seurat@meta.data$Cluster_ID)
# This shows you a table of nuclei counts per cell type and experimental group.

DE_female_aging_sig <- DE_female_aging[which(DE_female_aging$p_val_adj 
                                             < 0.05), ] 
# This subsets the data to show only significant DEGs.

table(DE_female_aging_sig$cell_type)
# This shows you the number of significant DEGs per cell type.

DE_female_aging_gene_freq <- as.data.frame(table(DE_female_aging_sig$gene))
# This shows you the DEGs most commonly affected across cell types.

library(hdWGCNA)
# This package can be used to run weighted correlation network analysis. Note
# that this analysis is computationally intensive and should ideally be done 
# via the GSAF core/POD R server. 
```