---
title: "snRNA-Seq Volcano Plots"
output: html_notebook
---
After running DESeq2/edgeR, first we will remove rows that have NAs.
```{r}
library(ggplot2)
library(ggrepel)
library(ggbreak)
library(dplyr)
library(tidyr)

DE_female_aging <- DE_female_aging %>% drop_na()
DE_male_aging <- DE_male_aging %>% drop_na()
DE_male_treatment <- DE_male_treatment %>% drop_na()
```
Next, we will create two rows to each data frame that will indicate whether
genes are significantly upregulated or downregulated. 
```{r}
DE_female_aging$DiffExpressed <- "NO"
DE_female_aging$DiffExpressed[DE_female_aging$avg_logFC > 0.25 & 
                                DE_female_aging$p_val_adj < 0.05] <- "UP"
DE_female_aging$DiffExpressed[DE_female_aging$avg_logFC < -0.25 & 
                                DE_female_aging$p_val_adj < 0.05] <- "DOWN"
DE_male_aging$DiffExpressed <- "NO"
DE_male_aging$DiffExpressed[DE_male_aging$avg_logFC > 0.25 & 
                                DE_male_aging$p_val_adj < 0.05] <- "UP"
DE_male_aging$DiffExpressed[DE_male_aging$avg_logFC < -0.25 & 
                                DE_male_aging$p_val_adj < 0.05] <- "DOWN"
DE_male_treatment$DiffExpressed <- "NO"
DE_male_treatment$DiffExpressed[DE_male_treatment$avg_logFC > 0.25 & 
                                DE_male_treatment$p_val_adj < 0.05] <- "UP"
DE_male_treatment$DiffExpressed[DE_male_treatment$avg_logFC < -0.25 & 
                                DE_male_treatment$p_val_adj < 0.05] <- "DOWN"
```
Volcano plots will then be created for astrocytes, microglia, and 
oligodendrocytes using the following code. Astrocytes below:
``` {r}
female_aging_astrocytes_top10 <- head(arrange(DE_female_aging
                                              [which(DE_female_aging$cell_type 
                                              == "Astrocytes"), ], 
                                              p_val_adj), 2) 
DE_female_aging$Label <- if_else(DE_female_aging$gene %in% 
                                female_aging_astrocytes_top10$gene, 
                                DE_female_aging$gene, NA)
FemaleAging_Astrocytes <- ggplot(data = DE_female_aging
                      [which(DE_female_aging$cell_type == "Astrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-7.5, 7.5) + ylim(0, 8) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Female - Astrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_female_aging$Label <- NA
rm(female_aging_astrocytes_top10)

male_aging_astrocytes_top10 <- head(arrange(DE_male_aging
                                              [which(DE_male_aging$cell_type 
                                              == "Astrocytes"), ], 
                                              p_val_adj), 10) 
DE_male_aging$Label <- if_else(DE_male_aging$gene %in% 
                                male_aging_astrocytes_top10$gene, 
                                DE_male_aging$gene, NA)
MaleAging_Astrocytes <- ggplot(data = DE_male_aging
                    [which(DE_male_aging$cell_type == "Astrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-7.5, 7.5) + ylim(0, 8) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male - Astrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_aging$Label <- NA
rm(male_aging_astrocytes_top10)

male_treatment_astrocytes_top10 <- head(arrange(DE_male_treatment
                                              [which(DE_male_treatment$cell_type 
                                              == "Astrocytes"), ], 
                                              p_val_adj), 10) 
DE_male_treatment$Label <- if_else(DE_male_treatment$gene %in% 
                                male_treatment_astrocytes_top10$gene, 
                                DE_male_treatment$gene, NA)
MaleTreatment_Astrocytes <- ggplot(data = DE_male_treatment
                    [which(DE_male_treatment$cell_type == "Astrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-7.5, 7.5) + ylim(0, 8) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male M. vaccae - Astrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_treatment$Label <- NA
rm(male_treatment_astrocytes_top10)

FemaleAging_Astrocytes
MaleAging_Astrocytes
MaleTreatment_Astrocytes

ggsave("Astrocytes - Female Aging.tiff", plot = FemaleAging_Astrocytes, 
       units = "in", width = 5, height = 4, dpi = 300)
ggsave("Astrocytes - Male Aging.tiff", plot = MaleAging_Astrocytes, 
       units = "in", width = 5, height = 4, dpi = 300)
ggsave("Astrocytes - Male Treatment.tiff", plot = MaleTreatment_Astrocytes, 
       units = "in", width = 5, height = 4, dpi = 300)
```
Microglia below:
```{r}
female_aging_microglia_top10 <- head(arrange(DE_female_aging
                                              [which(DE_female_aging$cell_type 
                                              == "Microglia"), ], 
                                              p_val_adj), 10) 
DE_female_aging$Label <- if_else(DE_female_aging$gene %in% 
                                female_aging_microglia_top10$gene, 
                                DE_female_aging$gene, NA)
FemaleAging_Microglia <- ggplot(data = DE_female_aging
                      [which(DE_female_aging$cell_type == "Microglia"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-10, 10) + ylim(0, 30) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Female - Microglia") +
  theme_classic() +
  theme(legend.position = "none")
DE_female_aging$Label <- NA
rm(female_aging_microglia_top10)

male_aging_microglia_top10 <- head(arrange(DE_male_aging
                                              [which(DE_male_aging$cell_type 
                                              == "Microglia"), ], 
                                              p_val_adj), 0) 
DE_male_aging$Label <- if_else(DE_male_aging$gene %in% 
                                male_aging_microglia_top10$gene, 
                                DE_male_aging$gene, NA)
MaleAging_Microglia <- ggplot(data = DE_male_aging
                    [which(DE_male_aging$cell_type == "Microglia"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("black", "blue", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-10, 10) + ylim(0, 30) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male - Microglia") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_aging$Label <- NA
rm(male_aging_microglia_top10)

male_treatment_microglia_top10 <- head(arrange(DE_male_treatment
                                              [which(DE_male_treatment$cell_type 
                                              == "Microglia"), ], p_val_adj), 3) 
DE_male_treatment$Label <- if_else(DE_male_treatment$gene %in% 
                                male_treatment_microglia_top10$gene, 
                                DE_male_treatment$gene, NA)
MaleTreatment_Microglia <- ggplot(data = DE_male_treatment
                    [which(DE_male_treatment$cell_type == "Microglia"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-10, 10) + ylim(0, 5) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male M. vaccae - Microglia") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_treatment$Label <- NA
rm(male_treatment_microglia_top10)

FemaleAging_Microglia
MaleAging_Microglia
MaleTreatment_Microglia

ggsave("Microglia - Female Aging.tiff", plot = FemaleAging_Microglia, 
       units = "in", width = 5, height = 4, dpi = 300)
ggsave("Microglia - Male Aging.tiff", plot = MaleAging_Microglia, 
       units = "in", width = 5, height = 4, dpi = 300)
ggsave("Microglia - Male Treatment.tiff", plot = MaleTreatment_Microglia, 
       units = "in", width = 5, height = 4, dpi = 300)
```
Oligodendrocytes below:
```{r}
female_aging_oligodendrocytes_top10 <- head(arrange(DE_female_aging
                                              [which(DE_female_aging$cell_type 
                                              == "Oligodendrocytes"), ], 
                                              p_val_adj), 10) 
DE_female_aging$Label <- if_else(DE_female_aging$gene %in% 
                                female_aging_oligodendrocytes_top10$gene, 
                                DE_female_aging$gene, NA)
FemaleAging_Oligodendrocytes <- ggplot(data = DE_female_aging
                      [which(DE_female_aging$cell_type == 
                               "Oligodendrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-8, 8) + ylim(0, 120) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Female - Oligodendrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_female_aging$Label <- NA
rm(female_aging_oligodendrocytes_top10)

male_aging_oligodendrocytes_top10 <- head(arrange(DE_male_aging
                                              [which(DE_male_aging$cell_type 
                                              == "Oligodendrocytes"), ], 
                                              p_val_adj), 1) 
DE_male_aging$Label <- if_else(DE_male_aging$gene %in% 
                                male_aging_oligodendrocytes_top10$gene, 
                                DE_male_aging$gene, NA)
MaleAging_Oligodendrocytes <- ggplot(data = DE_male_aging
                    [which(DE_male_aging$cell_type == "Oligodendrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-8, 8) + ylim(0, 120) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male - Oligodendrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_aging$Label <- NA
rm(male_aging_oligodendrocytes_top10)

male_treatment_oligodendrocytes_top10 <- head(arrange(DE_male_treatment
                                              [which(DE_male_treatment$cell_type 
                                              == "Oligodendrocytes"), ], 
                                              p_val_adj), 1) 
DE_male_treatment$Label <- if_else(DE_male_treatment$gene %in% 
                                male_treatment_oligodendrocytes_top10$gene, 
                                DE_male_treatment$gene, NA)
MaleTreatment_Oligodendrocytes <- ggplot(data = DE_male_treatment
                    [which(DE_male_treatment$cell_type == 
                             "Oligodendrocytes"), ], 
                      aes(x = avg_logFC, y = -log10(p_val_adj), 
                          col = DiffExpressed, label = Label)) +
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="red") +
  geom_hline(yintercept=-log10(0.05), col="red") +
  labs(x=expression(Log[2]~Fold~Change)) + 
  labs(y=expression(-Log[10](p~Value))) + 
  xlim(-8, 8) + ylim(0, 6) +
  geom_point() +
  geom_text_repel() +
  ggtitle("Male M. vaccae - Oligodendrocytes") +
  theme_classic() +
  theme(legend.position = "none")
DE_male_treatment$Label <- NA
rm(male_treatment_oligodendrocytes_top10)

FemaleAging_Oligodendrocytes
MaleAging_Oligodendrocytes
MaleTreatment_Oligodendrocytes

ggsave("Oligodendrocytes - Female Aging.tiff", 
       plot = FemaleAging_Oligodendrocytes, units = "in", width = 5, height = 4,
       dpi = 300)
ggsave("Oligodendrocytes - Male Aging.tiff", 
       plot = MaleAging_Oligodendrocytes, units = "in", width = 5, height = 4,
       dpi = 300)
ggsave("Oligodendrocytes - Male Treatment.tiff", 
       plot = MaleTreatment_Oligodendrocytes, units = "in", width = 5, 
       height = 4, dpi = 300)
```